Asymmetric relationships and their effects on coexistence

### 2023-04-26 paper/Paper.Rnw
working on (hopefully) final edits

%Intro to storage effects and why they might be influenced by ATAs
It is possible to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
may be relevant to them. Storage effects allow competing species to fluctuate in abundance without any going extinct
by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
Storage effects hinge on positive covariation between the fluctuating quality of the environment
for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
and on the assumption that optimal environmental conditions for distinct species are also distinct,
so that no two species find the same conditions to be optimal. Then, good environmental
conditions for a currently common species are paired with strong intraspecific competition, because of
positive EC covariance and the commonness of the species. But good environmental conditions for 
a currently rare species are not as closely accompanied by strong competition because the rarity of
the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model, which was originally introduced as a model of coral reef fish competing for a set of $N$ breeding territories, 
is as follows.
Letting $N_i(t)$ denote the adult population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is an adult mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
Notation for the whole paper is summarized in Table \ref{tab:notation}.
For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
distributed (iid) through time. In three distinct similations (Fig. \ref{fig:pedagog1}d-f), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).

%Goal para
We previously demonstrated that ATAs are common in nature can influence ecological phenomena
\citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}; and our arguments above
indicate that ATAs may influence species coexistence. Thus, we here seek to quantify the consequences 
of ATAs for competition, coexistence, and the maintenance of biodiversity, by means of the 
following specific goals. 
G1) To augment MCT to show formally how ATAs play a role in coexistence. 
MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATAs contribute to these components. 
We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
spatial variation can be considered in future work.
G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
sometimes coexist with ATAs but not without, or vice versa. This is to help 
illuminate whether ATA contributions to coexistence and biodiversity maintenance
are likely to be negligible or meaningfully large. 
G3) To provide an empirical example of species coexistence which highlights the role ATAs can play and the
potential importance of ATAs relative to other mechanisms of coexistence.
The example is a diatom system in a chemostat. In the Discussion, we consider whether climate change may alter
ATAs of environmental variables and thereby influence competition and coexistence.
Overall, our study presents a new mechanism of species coexistence
and a means of understanding its theoretical and empirical importance.

@@

%Two-parameter plots
%***DAN: This para will likely have to be revised depending on how Jasmin revises the diatom figure presentation
We also generated two-dimensional plots, where
two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed
at the value used by \cite{Descamps_2005}. For the two-dimensional plots, the quantity  
$\Delta_i^{[EC]}/\text{GWR}$ was displayed using color, to convey the 
importance of the ATA contribution relative to the whole GWR. Larger magnitudes
of this quantity indicate greater importance of ATAs. The case
$\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/\text{GWR}>1$)
corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{GWR}<0$
(which also produces $\Delta_i^{[EC]}/\text{GWR}>1$) corresponds to ATA exclusion.
Thus the condition $\Delta_i^{[EC]}/\text{GWR}>1$ encompasses the cases for which ATA
exclusion or rescue occurred, though the condition does not delineate, by itself,
which of these was the case.

All computations were done in R on an ordinary laptop running Ubuntu Linux. 
Complete computer codes for the project are stored at
\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.

@@

%***DAN: check this again after Jasmin gets done making the new version of Fig. 5
Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
show that ATA contributions to coexistence often have magnitudes comparable to the 
magnitudes of other terms of the extended MCT decomposition, and that 
ATA rescue occurs for some combinations of parameters for this system (Fig. \ref{fig:fig5}a-c).

%***DAN: finish it once Jasmin finishes d-f
Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f show that...

@@

%***DAN: Jasmin, please add to the list anyone you talked to at UO about this, and anyone who read it and commented 
%for you. Be generous here - even short convos merit an acknowlegement.
The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Laurn Hallett, and 
Robin Snyder for helpful discussions. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the devlopment of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig1_a-h.pdf}
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (g,h)]. 
The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
pictured and then transforming the marginals as specified in Methods.
(d-f) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
respectively corresponding variables (a-c; see Introduction).
(g,h) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (g) showing 
left-tail association and (h) showing right-tail association. \label{fig:pedagog1}}
\end{figure}

@@

%***DAN: Jasmin, this figure looks really great. Here are a few minor things I would like to ask you to
%do to further optimize:
%1) Currently there are axis labels and numbers only on the bottom and left panels, which makes sense.
%But the axis tick marks needs to occur on all panels. I noticed that this would be an imporvement
%when I found myself wanting to see where the ATA line was between the x axis and the GWR line, but it was 
%a little hard to see where the x axis was becuase there were no tick marks. It is possible to have tick marks
%but no numbers of labels in R, and that's what we need here.
%2) The font of the x axis label, "log fcundity standard deviation, \sigma" seems slightly different from the 
%other textual labels. Can you please make it consistent? Perhaps it is not bold and the others are bold, or 
%something.
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/decompFigLLN.pdf}
\caption{The contribution of ATAs to species coexistence was sometimes, but not always, 
meaningfully large compared to 
other components of the MCT decomposition for the log-normal fecundities lottery model.
All panels show GWR and the decomposition of
GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
$\delta$ and $\mu_1-\mu_2$ 
are fixed at values specified in the margins, and GWR and components of the decomposition
are plotted against $\sigma$. The contribution of ATAs ($\Delta_i^{[EC]}$) sometimes, but not always, 
has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
to species coexistence or failure to coexist. In some instances, the GWR line falls 
between the ATA line and the $x$ axis, indicating that, for those parameter values, ATAs 
make the difference between species
coexistence and exclusion of the weaker competitor. These instances, which are shaded, correspond to
ATA rescue or exclusion (Theory). Standard errors of plotted quantities 
were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) uses 
$q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$; though results differed
in detail, main conclusions were the same. \label{fig:fig3}}
\end{figure}

@@

%***DAN: Jasmin, can you please ensure that ATA exclusion and rescue are 
%shaded and keyed in the legend for panels a-c in the same way as the earlier figures?
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{ATA effects and ATA rescue and exclusion in the diatom system. 
The contribution of ATAs to the GWR of \emph{F. crotonensis}
in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
other components of the MCT decomposition (a-c). Panels a-c show GWR and the decomposition of
GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
For each panel, 
two of the three parameters defining the temperature oscillation (the amplitude, $a$; period, $P$; 
and mean, $\theta_0$; see Methods) were fixed at the values used by 
\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C), and 
GWR and components of the MCT decomposition
were plotted against the third parameter. The original values of the varying parameters 
are indicated by vertical dotted lines. 
%***DAN: Change the above sentence if Jasmin changes the font of the line
ATA contributions ($\Delta_i^{[EC]}$)
had magnitude comparable to other components, 
indicating that ATAs can contribute meaningfully
to diatom coexistence in this system. In some instances, the GWR line falls 
between the ATA line
and the horizontal axis, indicating that, for those parameter values, ATA rescue occurs. 
Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, 
was always positive both including and
excluding the effects of ATAs, so GWR for \emph{F. crotonensis}, decomposed here, is a 
sufficient coexistence metric.
%***DAN: Once Jasmin is done with the other part of the figure (the part that used to be Fig. 6), 
%finish this caption. Use the below caption, currently commented, to help. 
}  \label{fig:fig5}
\end{figure}

>>>>

%Intro to storage effects and why they might be influenced by ATAs
It is possible to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
may be relevant to them. Storage effects allow competing species to fluctuate in abundance, without any going extinct,
by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
Storage effects hinge on positive covariation between the fluctuating quality of the environment
for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
and on the assumption that optimal environmental conditions for distinct species are also distinct,
so that no two species find the same conditions to be optimal. Then, good environmental
conditions for a currently common species are paired with strong intraspecific competition, because of
positive EC covariance and the commonness of the species. But good environmental conditions for 
a currently rare species are not as closely accompanied by strong competition because the rarity of
the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model, which was originally introduced as a model of coral reef fish competing for a set of $N$ breeding territories, 
is as follows.
Letting $N_i(t)$ denote the adult population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is an adult mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
Notation for the whole paper is summarized in Table \ref{tab:notation}.
We assume for simplicity that the random variables $(B_1(t),B_2(t))$ are independent and identically 
distributed (iid) through time. In three distinct simulations (Fig. \ref{fig:pedagog1}d-f), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symmetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all 
simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).

%Goal para
We previously demonstrated that ATAs are common in nature can influence ecological phenomena
\citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}; and our arguments above
indicate that ATAs may influence species coexistence. Thus, we here seek to quantify the consequences 
of ATAs for competition, coexistence, and the maintenance of biodiversity, by means of the 
following specific goals. 
G1) We will augment MCT to show formally how ATAs play a role in coexistence. 
MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATAs contribute to these components. 
We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
spatial variation can be considered in future work.
G2) We will explore, using the simple lottery model, the relative magnitudes 
of the contributions to coexistence that come
from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
sometimes coexist with ATAs but not without, or vice versa. This is to help 
illuminate whether ATA contributions to coexistence and biodiversity maintenance
are likely to be negligible or meaningfully large. 
G3) We will provide an empirical example of species coexistence which highlights the role ATAs can play and the
potential importance of ATAs relative to other mechanisms of coexistence.
The example is a diatom system in a chemostat. In the Discussion, we consider whether climate change may alter
ATAs of environmental variables and thereby influence competition and coexistence.
Overall, our study presents a new mechanism of species coexistence
and a means of understanding its theoretical and empirical importance.

%***DAN: Dan made it to here on 2022 04 26

@@

%Two-parameter plots
We also generated two-dimensional plots, where
two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed
at the value used by \cite{Descamps_2005}. For the two-dimensional plots, the quantity  
$\Delta_i^{[EC]}/|\text{GWR}|$ was displayed using color, to convey the 
importance of the ATA contribution relative to the whole GWR. Larger magnitudes
of this quantity indicate greater importance of ATAs. The case
$\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/|\text{GWR}|>1$)
corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{GWR}<0$
(which produces $\Delta_i^{[EC]}/|\text{GWR}| < -1$) corresponds to ATA exclusion.
Thus the value of $\Delta_i^{[EC]}/|\text{GWR}|$ indicates the importance of ATAs
and also whether ATA rescue or exclusion occurred.

All computations were done in R on a laptop running Ubuntu Linux. 
Complete computer codes for the project are stored at (link to be added).
%\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.

@@

Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
show that ATA contributions to coexistence often have magnitudes comparable to the 
magnitudes of other terms of the extended MCT decomposition, and that 
ATA rescue occurs for some combinations of parameters for this system.

Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f reinforce the 
same conclusions. The plotted quantity $\Delta_i^{[EC]}/|\text{GWR}|$ was often large,
indicating that ATAs were often a substantial contributor to whether the diatom species 
coexist. For instance, well over half the area of Fig. \ref{fig:fig5}f showed values of 
$\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1/4$, with large portions of
Fig. \ref{fig:fig5}d-e satisfying the same criterion. Substantial portions of 
panels d-f also show values of $\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1$,
corresponding to ATA rescue. ATA exclusion did not occur for this system for the 
parameter ranges we considered.

@@

%***DAN: Jasmin, please add to the list anyone you talked to at UO about this, and anyone who read it and commented 
%for you. Be generous here - even short convos merit an acknowlegement.
The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Laurn Hallett, and 
Robin Snyder for helpful discussions. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

@@

%***DAN: Jasmin, could you, with reasonable effort, move panels around? Specifically, put the panels which
%are now g ang h in the middle instead of on the right (and they will then be called panels d and e) and 
%put what are now panels d-f on the right (and thwill then be called f-h)? The reason is, we refer to
%g-h before we refer to d-f, and the reader may become confused by this, including possibly havig trouble 
%finding the parts of the caption that refer to g-h while getting confused about the parts of the caption 
%that refer to d-e, which are about something that has not even been introduced at the point at which the 
%reader first is referred th panels g-h. 
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig1_a-h.pdf}
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (g,h)]. 
The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
pictured and then transforming the marginals as specified in Methods.
(d-f) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
respectively corresponding variables (a-c; see Introduction).
(g,h) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (g) showing 
left-tail association and (h) showing right-tail association. \label{fig:pedagog1}}
\end{figure}

@@

%***DAN: Jasmin, this figure looks really great. Here are a few minor things I would like to ask you to
%do to further optimize:
%1) Currently there are axis labels and numbers only on the bottom and left panels, which makes sense.
%But the axis tick marks needs to occur on all panels. I noticed that this would be an imporvement
%when I found myself wanting to see where the ATA line was between the x axis and the GWR line, but it was 
%a little hard to see where the x axis was becuase there were no tick marks. It is possible to have tick marks
%but no numbers of labels in R, and that's what we need here.
%2) The font of the x axis label, "log fecundity standard deviation, \sigma" seems slightly different from the 
%other textual labels. Can you please make it consistent? Perhaps it is not bold and the others are bold, or 
%something.
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/decompFigLLN.pdf}
\caption{The contribution of ATAs to species coexistence was sometimes, but not always, 
meaningfully large compared to 
other components of the MCT decomposition for the log-normal fecundities lottery model.
All panels show GWR and the decomposition of
GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
$\delta$ and $\mu_1-\mu_2$ 
are fixed at values specified in the margins, and GWR and components of the decomposition
are plotted against $\sigma$. The contribution of ATAs ($\Delta_i^{[EC]}$) sometimes, but not always, 
has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
to species coexistence or failure to coexist. In some instances, the GWR line falls 
between the ATA line and the $x$ axis, indicating that, for those parameter values, ATAs 
make the difference between species
coexistence and exclusion of the weaker competitor. These instances, which are shaded, correspond to
ATA rescue or exclusion (Theory). Standard errors of plotted quantities 
were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) uses 
$q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$; though results differed
in detail, main conclusions were the same. \label{fig:fig3}}
\end{figure}

@@

\begin{figure}
%\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{ATA effects and ATA rescue in the diatom system. 
The contribution of ATAs to the GWR of \emph{F. crotonensis}
in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
other components of the MCT decomposition (a-c). Panels a-c show GWR and the decomposition of
GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
For each panel, 
two of the three parameters defining the temperature oscillation (the amplitude, $a$; period, $P$; 
and mean, $\theta_0$; see Methods) were fixed at the values used by 
\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C), and 
GWR and components of the MCT decomposition
were plotted against the third parameter. The original values of the varying parameters 
are indicated by vertical dotted lines. 
ATA contributions ($\Delta_i^{[EC]}$)
had magnitude comparable to other components, 
indicating that ATAs can contribute meaningfully
to coexistence in this system. In some instances, the GWR line falls 
between the ATA line
and the horizontal axis, indicating that, for those parameter values, ATA rescue occurs. 
Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
Panels d-f show bivariate plots (Methods) for which the third parameter was 
fixed at the values used by \cite{Descamps_2005}. ATAs are again often an important factor helping
determine coexistence or exclusion, and ATA rescue happened for an appreciable fraction of the
explored parameter space.
For the parameters we considered, GWR of the other diatom species in the system, 
\emph{C. pseudostelligera}, was always positive both including and
excluding the effects of ATAs, so GWR for \emph{F. crotonensis}, decomposed here, is a 
sufficient coexistence metric.
}  \label{fig:fig5}
\end{figure}
